Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends

### 2021-11-15 Paper_online_Learning.tex
update paper

The ongoing COVID-19 pandemic has brought the world to a standstill. Many sectors have been affected ever since the outbreak of COVID 19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020. With the start of the pandemic a near total closure of schools were observed all across the globe. Africa and Oceania are in better positions in comparison to other regions with their increasing trend of fully open schools. In line with that, the Sri Lankan government also ordered to close schools from 12 March 2020 and its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances \autocite{wikieducation}.

The speed of these closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders. Sudden unexpected movement from classrooms to homes at large scale made students, teachers and parents vulnerable. The parents with limited education and resources; and students who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services. Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic \autocite{drane2020impact,daniel2020education,unescoadverse2020}.

Eventhough distance education has a long history that goes back to almost two centuries \autocite{spector2014handbook} with significant modifications,
alterations, and addition in the process of delivery and communication \autocite{moore2011learning}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously. Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to unexpected massive shutdown worldwide. Lack of experience, lack of awareness, lack of readiness for distance education put various stakeholders in the education system in panic, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic Figures \ref{fig:distanceLearningWorldAnalysis}. The search spikes coincide with increasing Covid-19 counts and related google trend searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.

@@

In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions. In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID 19 pandemic on education? (2) What solutions are in place to meet the need of distance education in terms of coverage, up-take, and learning during COVID 19 pandemic? (3) Which solutions have a wide attention and public interest both at the national and global level?

Confusion and stress for teachers: When schools close, especially unexpectedly and for unknown duration, teachers are often unsure of their obligations and how to maintain connections with students to support learning. Transitions to distance learning platforms tend to be messy and frustrating, even in the best circumstances. Limited time, and lack of knowledge about the best online tools and techniques available are other major challenges teachers encounter during the pandemic. In many contexts, school closures lead to furloughs or separations for teachers.\\
Parents unprepared for distance and home schooling: When schools close, parents are often asked to facilitate the learning of children at home and can struggle to perform this task. This is especially true for parents with limited education and resources.
Challenges creating, maintaining, and improving distance learning: Demand for distance learning skyrockets when schools close and often overwhelms existing portals to remote education. Moving learning from classrooms to homes at scale and in a hurry presents enormous challenges, both human and technical.

The design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content. It is important to know how the learning environment is used, and the influences of the tools and techniques that distinguish the differences in learning outcomes as the technology evolves \autocite{moore2011learning}.

As learning technology and its associated fields continue to evolve, practitioners and researchers have yet to agree on common definitions and terminologies (Lowenthal \& Wilson, 2010; Volery \& Lord, 2000) \autocite{moore2011learning}. There are many different distance learning solutions. Identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features. The high popularity or increasing online attention towards a product can indicate its prominence in the market, its usability and its impact. Further, developers also tend to improve the their products and serivices in response to increasing online search \autocite{willis2020using}

A big question for faculty members designing new online courses concerns their role as a teacher: what do I do when I teach online? Designing the course materials, interacting with students and giving them feedback, and assessing student work are obvious tasks for the online teacher. What is not
obvious, however, is how those tasks are carried out.

We primarily analyze quantitative digital footprint data on the Internet from January 2020 to 11.03.2020, including their representativeness.The google trend footprint will allow the teachers to narrow down thier search space and select prominent distance education solutions for their teaching purposes.

The remainder of this paper is organised as follows. Section 2 presents the related work to lay the foundation for the stray algorithm. Section 3 describes the limitations of the HDoutliers algorithm that hinder its performance. Section 4 presents the improved algorithm, stray, that addresses the limitations of the HDoutliers algorithm. Section 5 presents a comprehensive evaluation, illustrating the key features of the stray algorithm. Section 6 includes an application of stray algorithm related to pedestrian behaviour in the city of Melbourne, Australia. Section 7 concludes the article and presents future research directions.

@@

>>>>

The ongoing COVID-19 pandemic has brought the world to a standstill. Many sectors have been affected ever since the outbreak of COVID 19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020. With the start of the pandemic a near total closure of schools were observed all around the world. In line with that, the Sri Lankan government also ordered to close schools from 12 March 2020 and its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances \autocite{wikieducation}. Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactWorld}).

Eventhough distance education has a long history that goes back to almost two centuries \autocite{spector2014handbook} with significant modifications,
alterations, and addition in the process of delivery and communication \autocite{moore2011learning}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously. Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}). Surprisingly, the search spikes of distance education related searches coincide with increasing Covid-19 counts and related internet searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.

@@

The speed of these closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders. The parents with limited education and resources; and students who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services. Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic \autocite{drane2020impact,daniel2020education,unescoadverse2020}. How to preparing course materials, How to interact with students, how to assess and evaluate student work, how to provide feedback are some major concerns they had when preparing for their first online teaching experience. Limited time, and lack of knowledge about the best online tools and techniques available were some of the major challenges teachers encountered during the online course preparation. This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home.

In response to this grey situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a list of distance learning solutions that can be used to facilitate student learning during the period of school closure. there are many different distance learning solutions. The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content \autocite{moore2011learning}. However, getting familiar with all these available distance solutions is not feasible or practical due to limited time availability and need of urgent responses.

The popularity of a product greatly influences consumer purchasing decisions can indicate its prominence in the market, its usability and its impact \autocite{ahn2006utilizing}. Further, according to \textcite{willis2020using}, developers also tend to improve the their products and services in response to increasing online search. Therefore, the Internet represents a great opportunity to learn about public attention towards a product or service. It's also a way to narrow down thier search space and thereby identify suitable tools and techniques to create an appropriate learning environment.

Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior \autocite{nuti2014use}. In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions. In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID 19 pandemic on education? (2) What solutions are in place to meet the need of distance education in terms of coverage, up-take, and learning during COVID 19 pandemic? (3) Which solutions have a wide attention and public interest both at the national and global level?
We primarily analyze quantitative digital footprint data on the Internet from December 2019 to August 2021.The google trend footprint will allow the teachers to narrow down thier search space and select prominent distance education solutions for their teaching purposes.

The remainder of this paper is organized as follows. Section 2 presents the related work to lay the foundation for the Google footprint analysis. Section 3 presents the methodology followed in the study. Section 4 includes Google footprint analysis. Section 5 concludes the article and presents future research directions.

@@

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to unexpected massive shutdown worldwide.

### 2021-11-15 main.Rmd
update paper

```{r covidImpactContinent, fig.cap = "Region wise tracking of COVID-19 caused school closures and re-openings", fig.height=8}
```

The ongoing COVID-19 pandemic has brought the world to a standstill.  Many sectors have been affected ever since the outbreak of COVID 19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world [@daniel2020education]. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020. With the start of the pandemic a near total closure of schools were observed all across the globe. Africa and Oceania are in better positions in comparison to other regions with their increasing trend of fully open schools.  In line with that, the Sri Lankan  government also ordered to close schools from 12 March 2020 and  its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances [@wikieducation].

The speed of these closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders. Sudden unexpected movement from classrooms to homes at large scale made  students, teachers and parents vulnerable.  The parents  with limited education and  resources;  and  students  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services. Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID 19 pandemic [@drane2020impact; @daniel2020education; @unescoadverse2020].

Eventhough distance education has a long history that goes back to almost two centuries [@spector2014handbook] with significant  modifications,
alterations, and addition in the process of delivery and communication  [@moore2011learning],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously. Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to  unexpected  massive shutdown worldwide. Lack of experience, lack of awareness, lack of readiness for distance education put various stakeholders in the education system in panic, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic Figures \ref{fig:distanceLearningWorldAnalysis}. The search spikes coincide with increasing Covid-19 counts and related google trend searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.

```{r distanceLearningWorldAnalysis, fig.cap= "Google trends footprint analysis", fig.height=6}
```

In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions. In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of  COVID 19 pandemic on education? (2) What solutions are in place to meet the need of distance education in terms of coverage, up-take, and learning during COVID 19 pandemic? (3) Which solutions have a wide attention and public interest both at the national and global level?

Confusion and stress for teachers: When schools close, especially unexpectedly and for unknown duration, teachers are often unsure of their obligations and how to maintain connections with students to support learning. Transitions to distance learning platforms tend to be messy and frustrating, even in the best circumstances. Limited time, and lack of knowledge about the best online tools and techniques available are other major challenges teachers encounter during the pandemic.  In many contexts, school closures lead to furloughs or separations for teachers.  
Parents unprepared for distance and home schooling: When schools close, parents are often asked to facilitate the learning of children at home and can struggle to perform this task. This is especially true for parents with limited education and resources.
Challenges creating, maintaining, and improving distance learning: Demand for distance learning skyrockets when schools close and often overwhelms existing portals to remote education. Moving learning from classrooms to homes at scale and in a hurry presents enormous challenges, both human and technical.

The design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content. It is important to know how the learning environment is used, and the influences of the tools and techniques that distinguish the differences in learning outcomes as the technology evolves [@moore2011learning].

As learning technology and its associated fields continue to evolve, practitioners and researchers have yet to agree on common definitions and terminologies (Lowenthal & Wilson, 2010; Volery & Lord, 2000) [@moore2011learning]. There are many different distance learning solutions. Identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.  The high popularity or increasing online attention towards a product can indicate its prominence in the market, its usability and its impact. Further,  developers also tend to improve the their products and serivices in response to increasing online search [@willis2020using]

A big question for faculty members designing new online courses concerns their role as a teacher: what do I do when I teach online? Designing the course materials, interacting with students and giving them feedback, and assessing student work are obvious tasks for the online teacher. What is not 
obvious, however, is how those tasks are carried out.

We primarily analyze quantitative digital footprint data on the Internet from January 2020 to 11.03.2020, including their representativeness.The google trend footprint will allow the teachers to narrow down thier search space and select  prominent distance education solutions for their teaching purposes.

The remainder of this paper is organised as follows. Section 2 presents the related work to lay the foundation for the stray algorithm. Section 3 describes the limitations of the HDoutliers algorithm that hinder its performance. Section 4 presents the improved algorithm, stray, that addresses the limitations of the HDoutliers algorithm. Section 5 presents a comprehensive evaluation, illustrating the key features of the stray algorithm. Section 6 includes an application of stray algorithm related to pedestrian behaviour in the city of Melbourne, Australia. Section 7 concludes the article and presents future research directions.

@@

>>>>

```{r covidImpactContinent, fig.cap = "Region wise tracking of COVID-19 caused school closures and re-openings", fig.height=6}
```

The ongoing COVID-19 pandemic has brought the world to a standstill.  Many sectors have been affected ever since the outbreak of COVID 19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world [@daniel2020education]. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020. With the start of the pandemic a near total closure of schools were observed all around the world. In line with that, the Sri Lankan  government also ordered to close schools from 12 March 2020 and  its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances [@wikieducation]. Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactWorld}).

Eventhough distance education has a long history that goes back to almost two centuries [@spector2014handbook] with significant  modifications,
alterations, and addition in the process of delivery and communication  [@moore2011learning],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.  Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}). Surprisingly, the search spikes of distance education related searches coincide with increasing Covid-19 counts and related internet searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.

```{r distanceLearningWorldAnalysis, fig.cap= "Google trends footprint analysis", fig.height= 8}
```

The speed of these closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders.   The parents  with limited education and  resources;  and  students  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services. Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID 19 pandemic [@drane2020impact; @daniel2020education; @unescoadverse2020]. How to preparing course materials, How to interact with students, how to  assess and evaluate student work, how to provide feedback are some major concerns they had when preparing for their first  online  teaching experience. Limited time, and lack of knowledge about the best online tools and techniques available were some of the major challenges teachers encountered during the online course preparation. This also put an extra burden on parents, specially with limited education and resources,  as they were  expected to facilitate the required learning environment at home.

In response to this grey situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a list of distance learning solutions that can be used to facilitate student learning during the period of school closure. there are many different distance learning solutions.  The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content [@moore2011learning]. However, getting familiar with all these available distance solutions  is  not feasible or practical due to limited time availability and need of urgent responses.

The popularity of a product greatly influences consumer purchasing decisions can indicate its prominence in the market, its usability and its impact  [@ahn2006utilizing]. Further, according to @willis2020using, developers also tend to improve the their products and services in response to increasing online search. Therefore, the Internet represents a great opportunity to learn about public attention towards a product or service.  It's also a way to narrow down thier search space and  thereby identify suitable tools and techniques to create an appropriate learning environment.

Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior [@nuti2014use]. In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions. In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of  COVID 19 pandemic on education? (2) What solutions are in place to meet the need of distance education in terms of coverage, up-take, and learning during COVID 19 pandemic? (3) Which solutions have a wide attention and public interest both at the national and global level? 
We primarily analyze quantitative digital footprint data on the Internet from December 2019 to August 2021.The google trend footprint will allow the teachers to narrow down thier search space and select  prominent distance education solutions for their teaching purposes.

The remainder of this paper is organized as follows. Section 2 presents the related work to lay the foundation for the Google footprint analysis. Section 3 presents the methodology followed in the study. Section 4 includes Google footprint analysis. Section 5 concludes the article and presents future research directions.

@@

## Analysis

Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to  unexpected  massive shutdown worldwide.
